---
title: "OP_analysis"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    theme: journal
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html")
options(repos = c(CRAN = "https://cran.rstudio.com"))
```

```{r include=FALSE}
library(tidyverse)
library(readxl)
library(filesstrings)
library(readr)
library(XLConnect)
library(magrittr)
library(rlist)
library(ggpmisc)
library(gt)
library(gtsummary)
library(kableExtra)
library(gridExtra)
library(grid)
library(XML)
```



# **Introduction to the Data**

The data used in this analysis is pulled from excel documents that are the output of a MatLab program. Data from the following excel documents are read into separate dataframes before being combined into 2 master dataframes, `wt_ko_sheets` and `wt_ko_bwaves` :

- OP_KO_scotopic_mod1
- b-trace_KO_scotopic
- OP_WT_scotopic_mod1
- b-trace_WT_scotopic

# **Loading, Cleaning, and Tidying the Data**

Setting path to Excel files as variables
```{r}
ko_path <- "data/OP_KO_scotopic_mod1.xlsx"
ko_bwave_path <- "/Users/ryanhecht/RStudio/OP_data/data/b-trace_KO_scotopic.xlsx"
  
wt_path <- "data/OP_WT_scotopic_mod1.xlsx"
wt_bwave_path <- "/Users/ryanhecht/RStudio/OP_data/data/b-trace_WT_scotopic.xlsx"
```

## Creating functions for reading sheets into single df and extracting header names

- Here, every sheet in the workbook is given an "id" (its sheet name) before they are concatenated into one large data frame
- The first sheet is ignored since it does not adhere to the format of the rest
- The first column of every sheet is removed since it contains redundant information
- Ideally, I will not hard code the column range so that this would not break if more columns are added

`mk_header` utlizes `xl_header1` and `xl_concat` to contruct a preliminary data frame by pulling header titles from two different rows within the excel file



```{r echo=T, results='false'}
#function for getting header information from continuous cells on the original excel spreadsheet
#from_sheet specifies which sheet to take the headers from in case there is one sheet where the header values are not representative of the rest
xl_header1 <- function(path, header_range, from_sheet) {
    header_names <- read_excel(path = path, range = header_range, sheet = from_sheet)
    names(header_names)
}

#function for dropping sheets from the workbook, setting range, concatenating remaining sheets into single df
xl_concat <- function(path, sheet_rm, range, col_names) {
  path %>% 
    excel_sheets() %>% 
    list.remove(sheet_rm) %>% 
    purrr::set_names() %>% 
    map_df(~ read_excel(path = path, sheet = .x, range = range, col_names = col_names), .id = "sheet")
}

mk_header <- function(xl_path, mouse_range, stat_range, data_range, rm_sheets) {
  
  #gather header names for mice and statistics
  mouse_names <- xl_header1(xl_path, mouse_range, 3)
  stat_names <- xl_header1(xl_path, stat_range, 3)
  final_header <- append(stat_names, mouse_names)
  
  
  #if last element of final header doesnt end in NA, duplicate last element
  #replace any mouse names that begin with "..." with NA
  
  final_header <- final_header %>% 
    str_replace("^\\.+", NA_character_)
  
  last_element <- tail(final_header,1)
  
  if(!is.na(last_element)){
    final_header <- append(final_header,last_element)
  }
  else {
    final_header <- final_header
  }
  
  #replace all NA values with the mouse name that preceeds it. Allows duplication for left and right eyes
  final_header <- final_header %>%
    tibble() %>%
    mutate(zoo::na.locf0(final_header)) %>%
    pull()
  
  xl_concat(xl_path, rm_sheets, data_range, final_header)
}


#building preliminary dataframes
ko_sheets <- mk_header(ko_path,"E1:U1", "B2:D2", "B2:V1822", rm_sheets = c(1,2))
ko_bwaves <- mk_header(ko_bwave_path,"E1:U1", "B2:D2", "B2:V1822", rm_sheets = c(1,2))

wt_sheets <- mk_header(wt_path, "E1:R1", "B2:D2", "B2:R1822", rm_sheets = c(1,2))
wt_bwaves <- mk_header(wt_bwave_path,"E1:R1", "B2:D2", "B2:R1822", rm_sheets = c(1,2))
```


## Handling multi-header layout and building the data frame

- When imported into a data frame, this data set has 2 headers
  + **Header 1**: Mouse code names (i.e. 051, 049, 061...)
  + **Header 2**: Time, Mean, SEM, and mouse eyes (Left vs Right)
  
`one_header` combines the two headers of the data set into one while merging "Left" and "Right" names with the appropriate mouse

- use this function to re-write the name of the df used previously (e.g. "ko_sheets", "wt_sheets")

```{r}
one_header <- function(df, xl_path, rm_sheets, data_range) {
  
  #get rid of the "..." in the mouse names
  final_header <- names(df) %>% 
    str_replace("\\..*", "")
  
  #accounts for the addition of the "sheet" column not in the excel file
  final_header <- final_header[-1]
  
  #setting up list for the loop ahead
  row1 <-as.list(df[1,])
  row1 <- row1[-1]
  ncols <- length(row1)
  
  #take the "left" and "right" designations from the second row and append them to the header name in first row
  for (n in 1:ncols){
  if(row1[n] == "Left" | row1[n] == "Right"){
    as.character(final_header, row1)
    final_header[n] <- paste0(final_header[n],sep = "_", row1[n])
    }
  }
  
  xl_concat(xl_path, rm_sheets, data_range, final_header)
}

#contructing final data frames (OP data and B wave traces)
ko_sheets <- one_header(ko_sheets, ko_path, c(1,2),"B2:V1822")
head(ko_sheets)
ko_bwaves <- one_header(ko_bwaves, ko_bwave_path, c(1,2),"B2:V1822")
head(ko_bwaves)

wt_sheets <- one_header(wt_sheets, wt_path, c(1,2),"B2:R1822" )
head(wt_sheets)
wt_bwaves <- one_header(wt_bwaves, wt_bwave_path, c(1,2),"B2:R1822" )
head(wt_sheets)
```

## Tidying the Data

`tidy_mouse` is a tidying function that performs the following operations:

1. **Remove** all repeat header rows with no data (inlude the word "time")
2. **Use** `pivot_longer` to gather the data and separate single headers into multiple variables

```{r}
tidy_mouse <- function(df) {
  
  df <- df %>% 
    filter(!Time == "Time")
  
  headers <- as.list(names(df))
  ncols <- ncol(df)
  start_col <- which(str_detect(headers, "_")) %>% 
  first()

  df <- df %>% 
    pivot_longer(cols = start_col:ncols, names_to = c("mouse", "eye"), values_to = "response", names_sep = "_") %>% 
    mutate(response = as.double(response),
           SEM = as.double(SEM),
           Mean = as.double(Mean),
           Time = as.integer(Time),
           eye = as.factor(eye),
           mouse = as.factor(mouse)
           )
  df
}

# Applying to all existing dataframes
ko_sheets <- tidy_mouse(ko_sheets)
ko_bwaves <- tidy_mouse(ko_bwaves)

wt_sheets <- tidy_mouse(wt_sheets)
wt_bwaves <- tidy_mouse(wt_bwaves)

```

## Combining Tidy Data Frames Before Plotting

**Using `bind_rows` to preserve all columns**

Combined OP data
```{r}
wt_ko_sheets <- bind_rows("ko_sheets" = ko_sheets, "wt_sheets" = wt_sheets, .id = "df_source")
wt_ko_sheets
```

Combined b-wave data
```{r}
wt_ko_bwaves <- bind_rows("ko_bwaves" = ko_bwaves, "wt_bwaves" = wt_bwaves, .id = "df_source")
wt_ko_bwaves
```


# **Building Function for Finding Local Maxima and Minima**

Employing the `stat_peak` function to detect local extrema. Excludes peaks and valleys that are within 10 peaks of the highest/lowest 

- The `mouse_grapher` function plots the local extrema on the OP data

  + The arguments `span` and `threshold` control the thresholding parameters of the extrema detection
```{r results='asis', warning=FALSE}
mouse_grapher <- function(df, intensity, ms_code, time_window = 500, span = 7, threshold = 0.65, result = "both") {
  
  #identify the source sheet
  source_sheet <- NULL
   if (ms_code %in% ko_sheets$mouse == TRUE) {
     source_sheet <- "KO"
   }
   if (ms_code %in% wt_sheets$mouse){
     source_sheet <- "WT"
   }
  
  #check to see if mouse has response values before plotting
  mouse_response <- df %>% 
    filter(sheet == intensity, mouse == ms_code) %>% 
    select(response)
  
  if (all(!is.na(mouse_response))) {
  
  #plot the data and show local extrema
   mouse_plot <- df %>% 
    filter(mouse == ms_code , sheet == intensity , Time <= time_window, !is.na(response)) %>% 
    ggplot(aes(x = Time, y = response)) +
    geom_line(aes(color = eye), alpha = 0.7) +
    suppressWarnings(stat_peaks(aes(shape = eye, color = eye),
                                span = span, geom = "point", ignore_threshold = threshold, na.rm = TRUE))+
    suppressWarnings(stat_peaks(aes(color = eye, shape = eye),
               na.rm = TRUE,
               geom = "text", hjust = -0.2, vjust = 0.5, 
               angle = 90, check_overlap = TRUE, response.label.fmt = "i",
               span = span, ignore_threshold = threshold,
               alpha = 0.7))+
     scale_y_continuous(expand = c(.2,.2))+
    facet_grid(~ eye, scales = "fixed", as.table = TRUE) +
    ggtitle(ms_code, subtitle = source_sheet)+
    theme_bw()+
     theme(plot.title = element_text(size = 15, face = "bold"),
           plot.subtitle = element_text(face = "italic"))
  
  #collect the extrema values in a table   
  OP_max <- layer_data(mouse_plot, i = 2L) %>% 
    select(x, y, shape) %>% 
    mutate(eye = case_when(
      shape == 16 ~ "Left",
      shape == 17 ~ "Right")) %>%
    select(-shape) %>%
    kable(
      caption = paste0("Mouse:", sep = " ", ms_code, "|", sep = " ", "Intensity: ", intensity, "|", sep = " ", "Source: ", source_sheet),
      align = "c") %>%
    row_spec(0, bold = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "condensed"),
                  full_width = T,
                  fixed_thead = T)

  if(result == "plot") {
    return(mouse_plot)
  }
  if(result == "table"){
    print(OP_max)
  }
  if(result == "both"){
    print(mouse_plot)
    OP_max
  }
  
  
  }
  
  else {
    #random blank plot with text
   ggplot(mtcars, aes(x= wt, y = mpg)) +  
      geom_blank() +
      theme(plot.background = element_blank(),
            axis.text = element_blank(), 
            panel.background = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank()) +
      annotate("text", x = 3.5, y = 23, label = "Mouse has no recorded values")
  }
}
```


# **Implementing `mouse_grapher` for each mouse and intensity**

## Loop that iterates through all Sheets and Mice

```{r results='asis', fig.width= 10, fig.height= 7.5}
#Real sheets
sheets <- as.character(unique(wt_ko_sheets$sheet))
mice <- as.character(unique(wt_ko_sheets$mouse))

#test sheets
  # sheets <- as.character(c(-2.12, 0.6))
  # mice <- as.character(c("050", "051"))



#for loop attempt 
for (i in sheets) {
  cat("  \n##", i, "{.tabset}", "  \n")
  for (j in mice) {
    cat("  \n###", "**Mouse: **", j, "  \n")
    p1 <- mouse_grapher(wt_ko_sheets, intensity = i , ms_code = j , span = 7, threshold = 0.575, time_window = 300, result = "plot")
    p2 <- mouse_grapher(wt_ko_bwaves, intensity = i, ms_code = j, result = "plot", time_window = 899)

    title1 <- textGrob("Oscillatory Potential", gp= gpar(fontface = "bold"))
    title2 <- textGrob("B-wave", gp= gpar(fontface = "bold"))
    grid.arrange(p1, p2,
             top = title1,
             bottom = title2)
    cat("\n")
    mouse_grapher(wt_ko_sheets, intensity = i , ms_code = j , span = 7, threshold = 0.575, time_window = 300, result = "table")
  }
}
```

## Practice Threshold

```{r, results = 'asis', fig.width= 10, fig.height= 7.5}
# p3 <- mouse_grapher(wt_ko_sheets, intensity = -2.12 , ms_code = "078" , span = 7, threshold = 0.575, time_window = 300, result = "plot")
# p4 <- mouse_grapher(wt_ko_bwaves, intensity = -2.12, ms_code = "078", result = "plot", time_window = 899)
# 
#     title1 <- textGrob("Oscillatory Potential", gp= gpar(fontface = "bold"))
#     title2 <- textGrob("B-wave", gp= gpar(fontface = "bold"))
#     grid.arrange(p3, p4,
#              top = title1,
#              bottom = title2)
# mouse_grapher(wt_ko_sheets, intensity = -2.12 , ms_code = "078" , span = 7, threshold = 0.575, time_window = 300, result = "table")
```







